An Effective Model of HL-60 Differentiation

### 2017-03-08 Paper_v2.tex
Working on Table 2

@@

\noindent\subsubsection*{Estimation of model parameters.}
We estimated parameters appearing in the mRNA and protein balances, and the abundance of polymerases and ribosomes, from estimates of transcription and translation rates,
the half-life of a typical mRNA and protein, and typical values for the copies per cell of RNA polymerase and ribosomes from literature (Table ZZ).
For the remaining parameters, e.g., the $W_{ij}$ appearing in the control laws,
or parameters appearing in the transfer functions $f_{dj}$, were estimated from the gene expression and signaling data sets discussed here.
The saturation constants $K_{X}$ and $K_{T}$ were adjusted so that gene expression and translation resulted in gene products on a biologically realistic concentration scale.
Lastly, we calculated the concentration for gene $G_{j}$ by assuming, on average, that a cell had two copies of each gene at any given time.
Thus, the bulk of our gene expression parameters were based directly upon literature values, and were not adjusted during model identification.
The values used for the characteristic transcription/translation parameters, degradation constants and
macromolecular copy number are given in the supplemental materials along with the specific formulas required to calculate all derived constants.

@@

>>>>

\subsubsection*{Effective cell cycle arrest model.}
We formulated an effective model of the fraction of cells undergoing ATRA-induced cell cycle arrest, $\mathcal{A}$ as:
\begin{equation}\label{eq:a-eqn}
	\frac{d\mathcal{A}}{dt} = \epsilon_{1}\left(\mathrm{E2F}\right)^{\epsilon_{2}}
\end{equation}where $\epsilon_{i}$ were unknown parameters, and $\mathrm{E2F}$ denotes the abundance of the E2F transcription factor.
The parameters $\epsilon_{1}$ and $\epsilon_{2}$ were estimated from experimental cell cycle distribution measurements, by discretizing and \texttt{log} transforming Eq. \eqref{eq:a-eqn}:
\begin{eqnarray}
\log\epsilon_{1}+\epsilon_{2}\log\left(\mathrm{E2F}_{1}\right) & = & \log{\dot{\mathcal{A}}_{21}} \\
\log\epsilon_{1}+\epsilon_{2}\log\left(\mathrm{E2F}_{2}\right) & = & \log{\dot{\mathcal{A}}_{32}}
\end{eqnarray}
where subscripts denote the experimental time point (1 = 24 hr, 2 = 48 hr, and 3 = 72 hr), $\mathrm{E2F}_{i}$ denotes the $\mathrm{E2F}$ abundance at time point $i$,
and $\dot{\mathcal{A}}_{ij}$ denotes the time derivative of the fraction of arrested cells estimated between time point $i$ and $j$ (estimated from data).
We calculated candidate values for $\epsilon_{j}$ by solving the log-transformed system of equations for 10 biological replicates.

@@

\noindent\subsubsection*{Estimation of model parameters.}
We estimated parameters appearing in the mRNA and protein balances, and the abundance of polymerases and ribosomes, from estimates of transcription and translation rates,
the half-life of a typical mRNA and protein, and typical values for the copies per cell of RNA polymerase and ribosomes from literature (Table \ref{tbl:model-parameters}).
For the remaining parameters, e.g., the $W_{ij}$ appearing in the control laws,
or parameters appearing in the transfer functions $f_{dj}$, were estimated from the gene expression and signaling data sets discussed here.
The saturation constants $K_{X}$ and $K_{T}$ were adjusted so that gene expression and translation resulted in gene products on a biologically realistic concentration scale.
Lastly, we calculated the concentration for gene $G_{j}$ by assuming, on average, that a cell had two copies of each gene at any given time.
Thus, the bulk of our gene expression parameters were based directly upon literature values, and were not adjusted during model identification.
The values used for the characteristic transcription/translation parameters, degradation constants and
macromolecular copy number are given in the supplemental materials along with the specific formulas required to calculate all derived constants.

@@

\begin{center}

	\tablefirsthead{
	\hline
	\rowcolor[gray]{0.9} Symbol & Description & Value & Units & Source  \\}
	\topcaption{Characteristic model parameters estimated from literature.}\label{tbl:model-parameters}
	\begin{scriptsize}
	\begin{supertabular}{c|l|l|l|l}
	\hline

	$R_{1}$ & RNA polymerase abundance & 75,000 & copies/cell & \cite{Jackson:2000aa,Zhao:2014aa} \\
	$R_{2}$ & Ribosome abundance & 1 x 10$^{6}$ & copies/cell & \cite{Freitas2004} \\
	$G_{i}$ & Characteristic gene abundance & 2 & copies/cell & this study\\
	$K_{X}$ & Saturation constant transcription & 4,600 & copies/cell & this study\\
	$K_{T}$ & Saturation constant translation & 100,000 & copies/cell & this study\\
	& & & & \\
	$t_{1/2,m}$ & characteristic mRNA half-life (transcription factor) & 2 & hr & \cite{Yang:2003aa} \\
	$t_{1/2,p}$ & characteristic protein half-life & 10 & hr & \cite{Doherty:2009aa} \\
	$\theta_{m,j}$ & characteristic mRNA degradation constant & 0.34 & hr$^{-1}$ & derived \\
	$\theta_{p,j}$ & characteristic protein degradation constant & 0.07 & hr$^{-1}$ & derived \\
	& & & & \\
	$t_{d}$ & HL-60 doubling time & 19.5 & hr & this study \\
	$\mu$ & growth rate & 0.035 & hr$^{-1}$ & derived \\
	$k_{d}$ & death rate & 0.10$\mu$ & hr$^{-1}$ & derived \\
	& & & & \\
	$e_{T}$ & elongation rate RNA polymerase & 6 & nt/s & \cite{Darzacq:2007aa} \\
	$e_{X}$ & elongation rate Ribosome & 5 & aa/s & \cite{Bostrom:1986aa} \\
	$L_{T,o}$ & characteristic gene length & 15,000 & nt & [REFHERE] \\
	$L_{X,o}$ & characteristic transcript length & 5,000 & nt & [REFHERE] \\
	$k_{T}$ & characteristic transcription rate & 1.44 & hr$^{-1}$ & derived \\
	$k_{X}$ & characteristic translation rate & 3.60 & hr$^{-1}$ & derived \\
	& & & & \\
	$D$	& Diameter of an HL-60 cell & 12.4 & $\mu$m$^{3}$ & \cite{Rosenbluth:2006aa} \\
	$f_{C}$ & cytoplasmic fraction & 0.51 & dimensionless & \cite{Rosenbluth:2006aa} \\

	\hline
	\end{supertabular}
	\end{scriptsize}

\end{center}
